TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-07-31 paper/draft.tex
finish metallicity commentary

This paper is organized as follows: in Section \ref{sec:data} we describe the technical specifications of the data from the \apogee\ and \gaia\ surveys, as well as previous studies of M dwarfs in \apogee.
Section \ref{sec:methods} describes our model implementation using \thecannon\ framework, and Section \ref{sec:sample_selection} describes the our sample selection and derivation of training parameters.
\color{gcolor}{{In Section \ref{sec:results} we present our experimental results, evaluate the predictive accuracy of our models, apply our model to a selected test sample of $>$10,000 sources, and examine the validity of our parameters against color-temperature relations and isochrone grids.
Finally, in Section \ref{sec:discussion} we discuss model performance, future improvements, and implications of our results in the context of present and upcoming stellar and exoplanetary surveys.}\color{black} \\

@@

@@

\begin{figure*}[]
\begin{center}
	\includegraphics[width=.32\linewidth]{figures/met_comp1.png}
	\includegraphics[width=.32\linewidth]{figures/met_comp2.png}
	\includegraphics[width=.32\linewidth]{figures/met_comp3.png} 
	\includegraphics[width=.32\linewidth]{figures/met_comp4.png}
	\includegraphics[width=.32\linewidth]{figures/met_comp5.png}
	\includegraphics[width=.32\linewidth]{figures/met_comp6.png} 
	\includegraphics[width=.32\linewidth]{figures/met_comp7.png}
	\includegraphics[width=.32\linewidth]{figures/met_comp8.png}
	\includegraphics[width=.32\linewidth]{figures/met_comp9.png} 
	\caption{A comparison of ASPCAP and Cannon metallicities for 8,335 test sample sources separated into temperature bins of 100K. We see the scatter and metal-poor bias of ASPCAP metallicities clearly increases at decreasing temperatures.} 
	\label{fig:met_comp}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
	\includegraphics[height=5cm]{figures/toomre.png}
	\includegraphics[height=5cm]{figures/met_hist.png}
	\caption{Toomre diagram showing the distribution of the \cannon\ metallicities (left). Dashed lines at $v_{tot}=70$\,km/s and 180\,km/s show roughly the separation of thin/thick/halo stars. Histogram (right) shows the metallicity distribution of thin and thick disk stars, with blue corresponding to $v_{tot}<70$\,km/s and green corresponding to $70<v_{tot}<180$\,km/s respectively. The average metallicities of the distribution of the thin disk stars are $\feh=0.00\pm0.17\,$dex, and the average metallicities of the thick disk stars are $\feh=-0.14\pm0.19\,$dex.}
	\label{fig:toomre}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{figures/wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. }
\end{center}
\end{figure}

@@

As an additional check of our metallicity consistency, we cross-matched our M dwarf final sample with the \citealt{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 230 of the \apogee\ M dwarfs to have binary pairs (XX FGK+M, XX M+M, and XX WD+M).

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding roughly to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources. Separating the two populations into separate histograms (also shown in Figure \ref{fig:toomre}), we find that thick disk sources are nearly 1-$\sigma$ more metal poor than thin disk sources with the average thin disk metallicity being near-solar $\feh=0.00\pm0.17\,$dex, and the average metallicities of the thick disk stars being $\feh=-0.14\pm0.19\,$dex.
By contrast metallicities from ASPCAP show no significant difference in the distribution with an average thin disk metallicity of $\mh=-0.16\pm0.16\,$dex and an average thick disk metallicity of $\mh=-0.23\pm0.17\,$dex.

>>>>

This paper is organized as follows: in Section \ref{sec:data} we describe the technical specifications of the data from the \apogee\ and \gaia\ surveys, as well as previous studies of M dwarfs in \apogee.
Section \ref{sec:methods} describes our model implementation using \thecannon\ framework, and Section \ref{sec:sample_selection} describes the our sample selection and derivation of training parameters.
In Section \ref{sec:results} we present our experimental results, evaluate the predictive accuracy of our models, apply our model to a selected test sample of $>$5,800 sources, and examine the validity of our parameters against color-temperature relations and metallicities of binary pairs.
Finally, in Section \ref{sec:discussion} we discuss model performance, future improvements, and implications of our results. \\

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{figures/wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. The overall scatter between the 15 metallicity pairs is 0.08\,dex.}
\end{center}
\end{figure}

@@

\begin{figure*}
\begin{center}
	\includegraphics[width=.3\linewidth]{figures/met_comp1.png}
	\includegraphics[width=.3\linewidth]{figures/met_comp2.png}
	\includegraphics[width=.3\linewidth]{figures/met_comp3.png} 
	\includegraphics[width=.3\linewidth]{figures/met_comp4.png}
	\includegraphics[width=.3\linewidth]{figures/met_comp5.png}
	\includegraphics[width=.3\linewidth]{figures/met_comp6.png} 
	\includegraphics[width=.3\linewidth]{figures/met_comp7.png}
	\includegraphics[width=.3\linewidth]{figures/met_comp8.png}
	\includegraphics[width=.3\linewidth]{figures/met_comp9.png} 
	\caption{A comparison of ASPCAP and Cannon metallicities for 8,335 test sample sources separated into temperature bins of 100K. We see the scatter and metal-poor bias of ASPCAP metallicities clearly increases at decreasing temperatures.} 
	\label{fig:met_comp}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
	\includegraphics[height=5cm]{figures/toomre.png}
	\includegraphics[height=5cm]{figures/met_hist.png}
	\caption{Toomre diagram showing the distribution of the \cannon\ metallicities (left). Dashed lines at $v_{tot}=70$\,km/s and 180\,km/s show roughly the separation of thin/thick/halo stars. Histogram (right) shows the metallicity distribution of thin and thick disk stars, with blue corresponding to $v_{tot}<70$\,km/s and green corresponding to $70<v_{tot}<180$\,km/s respectively. The average metallicity of the thin disk distribution is $\feh=0.00\pm0.17\,$dex, and the average metallicity of the thick disk distribution is $\feh=-0.14\pm0.19\,$dex.}
	\label{fig:toomre}
\end{center}
\end{figure*}

@@

As a check of our test sample metallicity reliability, we cross-matched our M dwarf final sample with the \citealt{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 216 of the \apogee\ M dwarfs to have binary pairs (46 FGK+M, 155 M+M, and 15 WD+M). Out of the 155 M+M pairs, 8 pairs contained both pairs in \apogee. Cross-matching the list of FGK+M dwarf companions with several catalogs/surveys with measured stellar metallicities, we found an additional 7 sources with FGK metallicities from LAMOST \citep{Zhao:2012} and APOGEE (ASPCAP). The metallicity measurements for the 15 M dwarf binaries and their companions are given in Table \ref{table:binaries}, and the overall scatter is $0.08\,$dex--an improvement over the scatter of ASPCAP metallicities which is $0.15\,$dex for these 15 sources.

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding roughly to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources. Separating the two populations into separate histograms (also shown in Figure \ref{fig:toomre}), we find that thick disk sources are nearly 1-$\sigma$ more metal poor than thin disk sources with the average thin disk metallicity being near-solar $\feh=0.00\pm0.17\,$dex, and the average metallicities of the thick disk stars being $\feh=-0.14\pm0.19\,$dex.
On the other hand, metallicities from ASPCAP show no significant difference in distribution, with an average thin disk metallicity of $\mh=-0.16\pm0.16\,$dex and an average thick disk metallicity of $\mh=-0.23\pm0.17\,$dex.

Figure \ref{fig:met_comp} shows that ASPCAP metallicities are systematically more metal-poor than \cannon\ metallicities. We furthermore find that the bias is temperature-dependent: at the hottest temperatures ($\teff>3600\,$K) ASPCAP and \cannon\ metallicities are consistent to a scatter of $0.05-0.06\,$dex and offset by an average of $-0.12-0.15\,$dex, while at the lowest temperatures ($\teff<3200\,$K) ASPCAP and \cannon\ are consistent to a scatter of $\sim-0.13\,$dex and offset by an average of $-0.3\,$dex.
